\documentclass{article}
\usepackage[utf8]{inputenc}
\title{Physics 2 for Engineers: PHYS1155}
\author{Javier Coindreau}

\newtheorem{definition}{Definition}
\newtheorem{theory}{Theory}
\newtheorem{example}{Example}
\newtheorem{derivation}{Derivation}

\usepackage{amssymb}
\usepackage{enumerate}

\begin{document}


\begin{titlepage}
	\maketitle
	\tableofcontents{}
\end{titlepage}
\section{Electric Potential}
\subsection{Electric Potential Energy}
\subsubsection{Movable charges in a field}

What happens to charges that are free to move? Do they accelerate at a constant rate?

We could use the known forces of electric fields to determine their acceleration over time, but that results in multiple differentials due to the changing distance $r$ as charges move. It results in two coupled differential equations that both depend on $r_1,r_2$.

We can, however, move to viewing forces through terms of energy to determine changes in state between points.

\begin{definition}
	A conservative force is a force whose work done depends only on the initial and final positions, and for which we can define a potential energy function $U$
	\begin{equation} U(x)=-\int{F \cdot d\vec{x}} \end{equation}
\end{definition}

From this, we can use electric potential to translate electric forces to mechanical work.

This works because electricity is a conservative force, as only movement along electric field lines do any work.

For electrostatics, the potential is defined as 
\begin{equation} U_e=-\int{\frac{q_1 q_2}{4\pi\epsilon_0 r^2} r\hat \cdot dr}, U_e=\frac{q_1 q_2}{4\pi\epsilon_0 r}=\frac{kq_1 q_2}{r}
\end{equation}

Can be positive or negative, and units are joules.

\begin{example}
	Two protons are separated by 1cm and not in motion relative to one another. They are not fixed in place.
	\begin{enumerate}[a)]
		\item What is the initial potential energy of the configuration?

			Plug in: $ \frac{k e^2}{r_1}=\frac{8.99*10^9*1.6*10^{-19}}{0.01}=2.3*10^{-26} J$
		\item How fast are they moving at time $t$?

			$E_f=E_j, \frac{ke^2}{r_1}+0=\frac{ke^2}{r_2}+0.5 m v^2$
			\\The potential energy here is used as the energy of the entire system. Because of that, we need to multiply the kinetic energy times 2 to become $\frac{ke^2}{r_1}+0=\frac{ke^2}{r_2}+m v^2$

			Solve:
			$v=\sqrt{(\frac{ke^2}{r_1}-\frac{ke^2}{r_2})\frac{1}{m}}=3.62 m/s$
		\item How much work is done to bring them back together?

			$W_{me}=-W_e=\Delta U_E=U_1-U_2=\frac{ke^2}{r_1}-\frac{ke_2}{r_2}=mv^2=2.18*10^{-26} J$
	\end{enumerate}
\end{example}
\subsubsection{Special case of uniform electric fields}
When electric fields are constant, we can simplify the potential energy to rely on the field itself.

\begin{equation} F=qE=constant \iff \Delta U_e= q E\Delta r \end{equation}

	$\Delta r$ is the distance through the electric field. Much like gravitational potential energy, you must set a reference point to solve problems involving constant fields. 

\subsection{Electric Potential}

Electric potential is a scalar quantity denoting the force applied on a test charge by an electric field.

In an electric field, a charge experiences a force of $F=q E$. At a potential $V$, a charge has potentila energy of $V=\int{E \cdot dr}$. If the electric field is constant, $V=Eh$.

As an analog, weather systems work well. Defining isobaric contours (equal pressure lines), we can determine the direction of the wind by tracing perpendicular lines from high to low pressure. The pressure has a scalar value, while the wind has a vector value. This correlates well to voltage, where voltage contour lines can be drawn and the electric field flows from high to low voltage lines, where voltage is a scalar field and electric fields are vector fields.
:w:
\subsection{Relationship among fields and potentials}

Voltage is electric potential energy per charge, voltage is an electric field applied over a distance, electric force is electric field per charge, and electric potential energy is electric force over a distance.

\section{Capacitance and Dielectrics}
\subsection{Capacitance}
\subsubsection{Definition}
How do we store solar energy for use at other time? Common methods are compressed air, dams, molten salt, concrete, and capacitors!\\
\begin{tabular}{l|l}
	Batteries & Capacitors \\
	\hline
	Long charge time & Charge quickly\\
	Slow discharge & Fast discharge \\
	Corrosive chemicals & Not corrosive \\
	Low energy density & High energy density \\
	Designed for long term storage & For short term storage \\
\end{tabular}

A capacitor can be seen as two parallel plates storing charge between them. Complex capacitors often seem like they're cylindrical, but internally they always resemble two parallel plates. The material inside is the dielectric, which is always some form of insulator.

Mathematically it is represented through the value of capacitance, 

\begin{equation} C=\frac{Q}{V} \end{equation}

with the capacitance value C named the Farad. 

\subsubsection{Calculating capacitance}

Calculating capacitance always follows the same process:

\begin{enumerate}
	\item Find the electric field.
	\item Find the voltage across the plates
	\item Find the capacitance.
\end{enumerate}

Since they are parallel plates, the electric field will almost always be the same, $E=\frac{\sigma}{\epsilon_0}y$. The voltage is also always similar. Solving the integral, the potential difference becomes $\frac{\sigma d}{\epsilon_0}$.

From these equations, we can get another definition of the capacitance, 
\begin{equation} C=\frac{Q}{V}=\frac{Q}{\frac{\sigma d}{\epsilon}}=\frac{Q}{\frac{Q d}{A \epsilon}}=\frac{A \epsilon}{d} \end{equation}
The key assumption here is that the field is uniform, which is only really usable if the plate area $A$ is much greater than the distance $d$.

\subsection{Capacitors in circuits}

In capacitors in series, each capacitor has the same charge $Q$, because the buildup of charges one one induces a voltage across the other of equal magnitude. These voltages across the capacitors depend on the capacitance.
\begin{equation} V_{ab}=V_1+V_2=\frac{Q_1}{C_1}+\frac{Q_2}{C_2}=Q(\frac{1}{C_1}+\frac{1}{C_2})\end{equation}
\begin{equation} V=\frac{Q}{C_{eq}} \end{equation}

Their total capacitances $C_{eq}$, will be equal to 

\begin{equation} \frac{1}{C_{eq}}=\frac{1}{C_1}+\frac{2}{C_2}... \end{equation}

In parallel, the capacitors always have the same voltage. There is no guarantee that the charges will be equal, however, since charges depend on the capacitances of their respective capacitors.

\begin{equation} Q_{eq}=Q_1+Q_2=C_1V_1+C_2V_2=(C_1+C_2)V \end{equation}
\begin{equation} Q_{eq}=C_{eq}V \end{equation}

Their total capacitances are equal to 

\begin{equation} C_{eq}=C_1+C_2 \end{equation}

\subsubsection{Energy stored in capacitors}

Given a triangle of charges with side length $a$, what is the potential energy stored in this configurations? We can take the sum of the work needed
to bring each charge into position, $W_n$, and sum them.

For the first, there are no charges, so $W_1=0$. The second has one charge, so it is $qV_1$. Finally, the third has 2 charges already, so it does $qV_2$. Therefore,
the total work done is $2qV_1+qV_2$

When storing energy in a capacitor, this effect happens. The more charges that are moved to the capacitor plates, the more energy is needed to move the next charges. When solving the integral, we get $\frac{1}{2} \frac{Q}{V}$ as the total energy.

To calculate energy stored in parallel OR series capacitors, we simply find the equivalent capacitance $C_{eq}$ and solve $0.5 C_{eq} V^2$.

\subsection{Dielectrics}

The dielectric is the material between the conducting plates in the capacitor. 
Different dielectrics are used for: 

\begin{itemize}
	\item Having varied dielectric breakdown values
	\item Storing different amounts of charge
	\item Increasing capaacitance

\end{itemize}

Capacitanc increases because the induced charge density on the dielectric decreases between plates.

How much is it reduced by? It depends on the dielectric constant of the material: $K$. With the dielectfic constant, we can get the electric charge in the dielectric as $\frac{E_{empty}}{K}$

\subsubsection{Capacitance with dielectrics}

With a dielectric, the capacitance is 
\begin{equation} C_{new}=KC \end{equation}

What happens to the capacitance formulas though?

With a dielectric all electrostatic equations work, as long as we replace

\begin{equation} \epsilon_0 \rightarrow K\epsilon_0, K\epsilon_0 = \epsilon \end{equation}

\section{Current, Resistance, and Electromotive Force}

\subsection{Current}

\subsubsection{The Story of electrical circuits}

In a complete loop of conducting material, a battery (with voltage) produces an electric field in the conductor. These free charges accelerate through the 
conductor, but hit stuff on the way and lose energy (resistance), reaccelerating after collision. The net result is a slow drift of charges (drift velocity)
in one direction (current).

Drift velocities are very slow ($1 mm/s$ to $1 cm/s$).

\subsubsection{Ohm's Law}

According to Ohm's Law, \begin{equation} E=\rho J \end{equation} where $J$ is the current density and $\rho$ is the resistivity (resistivity is a material property, NOT RESISTANCE). 
$J$ is defined as \begin{equation}nqv_d=\frac{I}{A} \end{equation}
A more important term here, however is the current $I$. This is defined as charge per unit time, \begin{equation} I=\frac{dQ}{dt} \end{equation}

A key term is conventional current, which is the movement of positive charge. In real life, the positive charges don't move, but that ended up being 
the custom historically, and it helps to not keep track of negative charges. Overall, it doesn't really matter.

\subsubsection{Derivation of current density}

How much charge passes a point in unit time? $n=\frac{charge count}{volume}$. We can find the charge moving through the conductor with $\delta Q=Nq=nVq=nA\delta x q$, where N is the amount of charges. 
Getting derivative over time results in $\frac{\delta Q}{\delta t}=nAq\frac{\delta x}{\delta t}$. Finally, we get current, $I=nAqv_d$

\subsubsection{Current in a wire}

\begin{example}
	If the current through a 0.511mm copper wire is defined as $I(t)=300mAe^{0.5s^{-1}t}=I_0e^{-\lambda t}$.\\
	How much charge passes through per unit time? Well, we can integrate it $\int_0^4.5{I_0e^{-\lambda t}}=\frac{1}{-\lambda}I_0e^{-\lambda t}$. Evaluating
	from 0 to 4.5, we get $\frac{I_0}{-\lambda}(e^{-\lambda*4.5}-1$. Plugging in, $\frac{300}{0.5}(1-e^{0.5*4.5})=537 mA s=537 mC$
\end{example}

\subsection{Circuits}

What happens in a circuit?

There is a potential difference delivered by a potential source. This potential source drives charges around a conductor in a closed loop, crossing through resistors and
losing energy in the process. The charge carrier returns having lost all energy.

\subsubsection{Terminal Voltage}

The electromotive force applies a certain amount of work on a unit of charge. Unlike the emf, the voltage is the work done to move a charge from the terminals of the 
battery. While the difference is very minor, they are technically different. The terminal voltage includes the internal resistance of the battery, while the electromotive 
force does not.

\subsubsection{Power}

How much work does it take to move $dq$ through $V$?
$$ dW= dq V $$

The battery must then supply power equal to $P=\frac{dW}{dt}=\frac{dq}{dt}V=IV$. From Ohm's law, we can see how much a resistance dissipates power as

\begin{equation} P=I^2R=\frac{V^2}{R}=IV \end{equation}

\subsection{RC Circuits}

\subsubsection{Series and parallel resistance}

In series resistance, series resistors are arranged so that theres a single current path, meaning that the current is the same across them. The voltage drop is equal to 
the sum of their voltage drops. Therefore, the resistances are equal to $$R_{eq}=R_1+R_2+...=\sum_i{R_i}.$$

In parallel, the voltage drop across each is equal but the resistances are equal to the sum of each individual resistor. The equivalent resistance is $$R_{eq}=\sum_i{\frac{1}{R_i}}$$

\subsubsection{Kirchhoff's Rules}

The two important rules are:

\begin{itemize}
	\item Junction Rule: At a junction, the current entering must be equal to the current exiting the junction: $$\sum{I_{in}}+\sum{I_{out}}=0$$

	\item Loop Rule: The sum of voltage difference in any loop is 0.
\end{itemize}

When analyzing a circuit, it is important to keep consistency with current direction. It doesn't matter in what direction the current goes, but keep it consistent and 
invert the current if it comes out negative.

\subsubsection{RC Circuits}

An important distinction with RC ciircuits (resistors and capacitors) is that unlike resistor circuits, RC circuits evolve with time.
There are two sepearate situation with RC circuits:
\begin{enumerate}
	\item When capacitors are initially uncharged, they begin acting as wires and then act as breaks in the circuit.
	\item When capacitors are initially charged, they initially act as batteries and then act as breaks.
\end{enumerate}

With initially uncharged capacitors (assuming a battery, resistor, and capacitor), Kirchoffs loop rule gives $$ee-iR-\frac{q}{C}=0$$

This can be rewritten as a differential $$\frac{dq}{dt}=\frac{e}{R}-\frac{q(t)}{RC}$$

SInce we know how charge builds on it, we can solve the differential equation to get 
\begin{equation} q(t)=Q_f(q-e^{t/RC}), Q_f=Ce \end{equation}

where $Q_f$ is the final charge on the capacitor.

RC circuits have an interesting connection. The time constant, $\tau$, ends up being $\tau=RC$ where $\tau$ is in seconds. At time $t=\tau, q(\tau)=0.63 Q_f$.
At any point through the circuit, the current of the circuit at a function of time is $$I_0e^{-t/\tau}$$

When capacitors start charged, Kirchoff's loop rule tells us that the current becomes time dependent. $$q(t)=Q_0(e^{-t/RC})$$

At any time, the current through the circuit behaves the same as the uncharged capacitor circuit, $$I_0e^{-t/\tau}=Q_0/RC e^{-t/\tau}$$

\section{Magnetism}

Magnetic fields can be treated similarly to electric fields, but they have a few key differences

\begin{itemize}
	\item Electric fields are described by Coulomb's law, while magnets are described by Lorenz force law
	\item Electric fields affect all charges, magnets affect moving charges.
	\item Magnets are in Teslas.
	\item Unlike electric fields, magnets have no monopoles
\end{itemize}

Due to convection, the core of the Earth is a large magnet as well. Currently, magnetic south is in the geographical North and magnetic nort in the Soutn. For reference, the strength of the magnetic field on the surface of the earth is 1 gauss.

\subsection{Magnetic Flux}

Magnetic flux defined just as magnetic flux was:
$$\Phi=\int{\vec{B}* d\vec{A}}$$

In a uniform field, this solves to
$$\Phi=B*A$$

This flux is used alot more often than electromagnetic flux.

A key fact is that north and south poles can never be separated. Therefore, for magnetism, Gauss's Law in a closed surface is always 0, as equal amounts of positive and negative flux must exist.

\subsection{Lorentz Force Law}

A magnetic field exerts a force on a moving charged particle,

$$F=q v \times B=qvB\sin{\theta}$$

applying a force perpendicular to the field $B$ and velocity. The direction can be figured out from the right hand rule.

Because magnetic force is perpendicular to the charged particle's velocity, it must undergo circular motion in a uniform field:
$$F=qvB=\frac{mv^2}{r}$$

This is the basis for mass spectrometry.

We can solve for the radius of curvature, $r=\frac{mv}{qB}$

\subsection{Charges in magnetic fields}

In the presence of a magnetic field, a current carrying conductor experiences a force of $$F_B=I \vec{L}\times\vec{B}$$

In this case, the vector form of the length follows the same direction as the charge.
If the wire is not a straight line, you can integrate each small section of the wire.

In a current loop, a magnetic field applies differing amounts of force based on the direction at that point in the loop. This generates a torque. The torque in that loop
can be found by $$\tau=\mu\times B$$ where $\mu$ is defined as the magnetic dipole moment, $$\mu=I\vec{A}$$. In a non changing field, the torque aligns $\mu$ with B.

The magnetic dipole equation is valid for any loop shape, and its directio is determined by the right hand rule. This also associates a potential memory with the dipole, 
$$U=-\mu\cdot B=-\mu B \cos{\theta}$$

\subsection{Magnetic Field Sources}
A moving particle produces a magnetic field in concentric circles around it 
with strength $$B=\frac{\mu_0}{4\pi} \frac{q\vec{v}\times \hat{r}}{r^2}$$

$\mu_0$, the magnetic permeability, is the measure of a material to produce a

magnetic field, $$\mu_0=4\pi * 10^-7 T$$.

Some simplifications for the field are:

 $$B=\frac{\mu_0}{4\pi} \frac{q\vec{v}\times \hat{r}}{r^2}=\\
 \frac{\mu_0}{4\pi} \frac{q\vec{v}\times \vec{r}}{r^3}=\\
 \frac{\mu_0}{4\pi} \frac{q\vec{v}\sin{\theta}}{r^2}$$

 \subsubsection{Law of Biot-Savart}
 A current produces a magnetic field similar in behavior to that produced by
 the charge.
 $$B=\frac{\mu_0}{4\pi} \int{\frac{I \vec{dL}\times\hat{r}}{r^2}}$$

 \subsubsection{Ampere's Law}
 Like with electric fields, Ampere's law can be used in cases of symmetry similarly to Gauss.

 $$\int{B\cdot dl}=\mu_0 I$$

 \begin{enumerate}
	\item Use symmetry to illustrate field lines around an object
	\item Draw an Amperian loop in the direction of the field lines
	\item Integrate along a field line witha constant field
	\item Determine the enclosed current
	\item Solve for B
 \end{enumerate}


\section{Induction}

\subsection{Faraday's Law of Induction}
This law describes the induced electromotive force caused by a changing magnetic flux.
$$\epsilon=\frac{-d\omega_B}{dt}$$

The negative sign indicates that the EMF opposes the change in magnetic flux.

This describes the induced voltage created by a flux. If the field is constant, the equations below are valid:

$$\epsilon=\frac{dB}{dt}A\cos{\theta}-B\frac{dA}{dt}\cos{theta}+BA\frac{d\theta}{dt}\sin{theta}$$

When multiple turns are included, the EMF gets multiplied by the amount of turns.

An isolated conducting rod becomes a source of EMF when moved through a field. In general, this emf can be found with 

$$\epsilon=\int{|\vec{v}\times\vec{B}|\cdot dl}$$

This emf is applied from the center to the two ends.

A changing field through a solid conductor introduces an emf which produces eddy currents, which are swirling currents in small loops.
These currents then create a force that oppose the motion of the conductor.

An example of use of eddy currents is metal detectors, where a large magnet makes a field in metallic objects in the field, producing 
an eddy current. This current can then be measured.

\subsection{Induced Electric Field}
As shown earlier, an emf is produced by a changing magnetic flux. What happens if there is no conductor present, however?

An electric field is produced!

$$\int{\vec{E}\cdot dl}=-\frac{d\Omega}{dt}$$

This is a different electric field, since it is not electrostatic. This means that:

\begin{enumerate}
	\item Non-conservative
	\item Does not have an electric potential associated
	\item Still exerts a force on a charge
\end{enumerate}

\begin{example}
	Consider a very long solenoid with a radius R and a current
	 $$I(t)=I_0\sin{\omega t}$$

	 What is the magnetic field in the solenoid and the electric field outside?

	 Magnetic: $B(t)=\mu_0 n I_0\sin{\omega t}$

	 Electric: $E=?, \int{E\dot dl}=-\frac{d\Omega}{dt}$

	 $E\int{dl}=E(2\pi r)=-\frac{d}{dt}(\pi R^2 B(t))$

	 $E=\frac{\pi R^2}{2\pi r} (\mu_0 n I_0 \sin{\omega t})'=\frac{\pi R^2}{2\pi r} (\omega \mu_0 n I_0 \cos{\omega t})$
 \end{example}

\subsection{Displacement Currents}

What happens around a parallel plate capacitor when they are charging?

There is a displacement current between plates that induces a magnetic field, $$I_d=\epsilon_0\frac{d\Omega}{dt}$$

Charge on capacitors: $q=CV=A\epsilon_0 \frac{V}{d}=\epsilon_0 EA=\epsilon_0 \Omega$

When charging: $i_c=\frac{dq}{dt}=\epsilon_0\frac{d\Omega}{dt}=i_{disp}$

\section{Waves}

Maxwell's Equations are:
\begin{itemize}
	\item Gauss's Law: $\int{E\cdot dA}=0$
	\item Faraday's Law: $\int{E\cdot dL}=\frac{d\Omega_B}{dt}$
	\item Ampere's Law: $\int{E\cdot dL}=\frac{d\Omega_E}{dt}$

\end{itemize}

The fact that Faraday's law incites a change in the magnetic field, affecting Ampere's law which does the same for electric fields, 
represents the effect of electromagnetic waves.

In simple harmonic motion to waves, we can find the position of any vertical type of wave as 
$y(t)=\cos{\omega t + \phi}$. Even though SHM requires 1 mass, mechanical
waves require a collection of masses. This still works, however, because the "object" moving is the wave itself.

To begin a mechanical wave, we'd need a disturbance in some medium with elements that interact and a force that tries to restore them.
For EM waves, however, all that is needed is an accelerating particle or an "accelerating" field. NO MEDIUM IS NEEDED.

At a point in space, the position  of an oscillating element can be described as undergoing SMH.

$$\frac{d^2y(x_0,t)}{dt^2}=-\omega^2y-(x_0,t)$$

In addition, at some point in time , the positions can also be described in those terms, 

$$\frac{d^2y(x_0,t_0)}{dx^2}=-k^2y-(x,t_0)$$

This number k is considered to be the "wave number" of this wave.

A periodic, sinusoidal, traveling wave can be described by both

$$y(x,t)=A\sin{(kx\pm \omega t)}$$
$$y(x,t)=A\cos{(kx\pm \omega t)}$$

Many waves can be described as both. For example, a sound wave can be seen as either a displacement wave (of the air molecules)
or a pressure wave. 

$$s(x,t)=s_{max}\cos{(kx\pm \omega t)}$$
$$\Delta p(x,t)=\Delta p_{max}\sin{(kx\pm \omega t)}$$

\subsection{Wave speed}

Since the waveform shape is constant ($kt\pm \omega t = constant$), we can define the speed as 
$$v_{wave}=\pm \frac{\omega}{k}=\frac{\lambda}{T}=\lambda f$$

The speed of mechanical waves are determined solely by the properties of the medium, and the frequency and wavelength simply
define the result.

$$v=\sqrt{\frac{restoring force}{intertial term}}$$

As a basic one, the speed of a wave on a string is $v=\sqrt{\frac{F_T}{\mu}}$ where $\mu$ is the linear mass density.
\end{document}
